\bf\mftitle

### 2011-02-17 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@

The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}. Process of export consists of two phases described in detail in later sections.
\begin{enumerate}
	\item Preprocessing.
	\item The export to a string representation itself.
\end{enumerate}
Method \code{start()} first creates an instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input. Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.\\

@@

As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
\begin{itemize}
	\item Decide which elements should be defined globally.
	\item Remove unused elements.
	\item Find the top level element.
	\item Find an instance of element by its name.
\end{itemize}

@@

Minimal number of occurrences of an element to define its type globally can be altered if preferences, as well. See \ref{subsection:preferences}.

@@

Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language. This subsection covers only XSD features that the exporter supports. For list of supported features, see TODO ref.\\

Element is defined by XSD element \code{element}, specifying its name and type. Let \code{xs} be XMLSchema namespace.

@@

	\item \code{complexType}. This type can contain XSD element \code{xs:sequence} or \code{xs:choice}. Each of these elements can contain definitions of elements, \code{xs:sequence}s and \code{xs:choice}s again.
	\begin{verbatim}
	<xs:element name="Person">
	  <xs:complexType>
	    <xs:sequence>
	      <xs:element name="Name" type="xs:string"/>
	      <xs:element name="Surname" type="xs:string"/>
	      <xs:choice>
	      ...
	      </xs:choice>
	    </xs:sequence>
	  </xs:complexType>
	</xs:element>
	\end{verbatim}
	An \emph{empty element} is defined as empty \code{complexType}.
	\begin{verbatim}
	<xs:element name="EmptyElement">
	  <xs:complexType>
	  </xs:complexType>
	</xs:element>
	\end{verbatim}
\end{itemize}

@@

Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}.

@@

\begin{verbatim}
public AbstractElementsExporter(final PreprocessingResult preprocessingResult,
                                final Indentator indentator)
\end{verbatim}

Names of parameters are self explanatory. Parameter \code{preprocessingResult} is a result of preprocessing. Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter. This instance has not to be empty. Output of exporter is appended at the end of text held by the \code{indentator}. This behaviour is convenient when chaining output of several elements exporters.\\

@@

\code{GlobalElementsExporter} retrieves global elements from the result of preprocessing and creates global definition of their types. These definitions are appended to the \code{indentator}. The \code{indentator} should be set to a level of indentation at which it is desired to append the global type definitions (typically, no indentation). After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.\\

If a global type contains another elements, these are processed as following.

\begin{itemize}
	\item If a contained element is global or of a built-in type, its type is simply referenced.
	\item If a contained element is not global nor of a build-in type, it is fully defined in a place of its occurrence inside the global type.
\end{itemize}

@@

Examples of definition of global types.

Type generated from element named "GlobalElement", containing elements "Text", which is string, and element "AnotherGlobalElement", which is another global element.

@@

\code{RootElementExporter} retrieves the root element from the result of preprocessing and creates its definition. This definition is also appended to the \code{indentator} (supplied to constructor) at a level of indentation, the \code{indentator} is set. Way of handling the root element's subelements depends on if these are global elements or not.\\

@@

Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence. For example, root element "Root" contains element "A", which contains two another string elements.

@@

Export of attributes is done along with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.

@@

\begin{itemize}
	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
	\item Number of spaces in output per one level of indentation.
	\item Global type name prefix. It is a string which will be inserted before a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
	\item Global type name suffix. It is a string which will be appended after a name of a type, which is derived from element's name. Can be also an empty string. (Only if generation of global elements is active.)
\end{itemize}

@@

This module is just basic implementation so many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.

@@

This problem appears if input grammar contains an alternation of two or more concatenations, which begin with at least one same element. Part of a problematic XSD is for example following.

\begin{verbatim}
...
  <xs:choice>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="B" type="TB"/>
    </xs:sequence>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="C" type="TC"/>
    </xs:sequence>
  </choice>
...
\end{verbatim}

@@

@@

Flow of data in this module is following.
\begin{enumerate}
	\item \code{Preprocessor} topologically sorts elements (rules) it got on input.
	\item For each element, its occurrence count is computed.
	\item Unused elements (occurrence count equals 0) are removed.
	\item For each element, it is determined if it type will be defined as a global type or not.
	\item For each global element, its type is exported to a XSD representation by \code{GlobalElementsExporter}.
	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{RootElementExporter}.
	\item String representation of the schema is returned along with the information that file extension should be "xsd".
\end{enumerate}

>>>>

For the whole document, let \code{xs} be XMLSchema namespace.

@@

The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provid\-er is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}. Process of export consists of two phases described in detail in later sections.
\begin{enumerate}
	\item Preprocessing.
	\item The export to a string representation itself.
\end{enumerate}
Method \code{start()} first creates an instance of \code{Preprocessor} class supplying the rules (elements) it got in the simplified grammar on input. Preprocessing itself is done by creating that instance (calling its constructor) and its results can be then retrieved by calling \code{getResult()} method. Purpose of preprocessing is to discover information necessary for export of elements. For example, determine elements which should be globally defined or which element is the root element.\\

@@

As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its is the following following.
\begin{itemize}
	\item Decide which elements should be defined globally.
	\item Remove unused elements.
	\item Find the top level element.
	\item Find an instance of element by its name.
\end{itemize}

@@

Minimal number of occurrences of an element to define its type globally can be altered if preferences as well. See \ref{subsection:preferences}.

@@

Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language. This subsection covers only XSD features that the exporter supports. For list of supported features, see \ref{subsection:supportedXSDFeatures}\\

Element is defined by XSD element \code{element}, specifying its name and type.

@@

	\item \code{complexType}. This type can contain XSD element \code{xs:sequence} or \code{xs:choice}. Each of these elements can contain definitions of elements, \code{xs:sequence}s and \code{xs:choice}s again.
	\begin{verbatim}
	<xs:element name="Person">
	  <xs:complexType>
	    <xs:sequence>
	      <xs:element name="Name" type="xs:string"/>
	      <xs:element name="Surname" type="xs:string"/>
	      <xs:choice>
	      ...
	      </xs:choice>
	    </xs:sequence>
	  </xs:complexType>
	</xs:element>
	\end{verbatim}
	An \emph{empty element} is defined as an empty \code{complexType}.
	\begin{verbatim}
	<xs:element name="EmptyElement">
	  <xs:complexType>
	  </xs:complexType>
	</xs:element>
	\end{verbatim}
\end{itemize}

@@

Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}. Whether an element is obligatory can determined from its metadata by the following code, assuming \code{attribute} is an instance of \code{Attribute} class.

\begin{verbatim}
if (attribute.getMetadata().containsKey(IGGUtils.REQUIRED)) { ...
\end{verbatim}

@@

\begin{verbatim}
public AbstractElementsExporter(PreprocessingResult preprocessingResult,
                                Indentator indentator)
\end{verbatim}

Names of parameters are self explanatory. Parameter \code{preprocessingResult} is a result of preprocessing. Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter. This instance doesn't need to be empty . Output of exporter is appended at the end of text held by the \code{indentator}. This behaviour is convenient when chaining output of several elements exporters.\\

@@

\paragraph{GlobalElementsExporter}
\code{GlobalElementsExporter} retrieves global elements from the result of preprocessing and creates global definition of their types. These definitions are appended to the \code{indentator}. The \code{indentator} should be set to a level of indentation at which it is desired to append the global type definitions (typically, no indentation). After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.\\

If a global type contains other elements, these are processed as following.

\begin{itemize}
	\item If a contained element is global or of a built-in type, its type is simply referenced.
	\item If a contained element is not global nor of a built-in type, it is fully defined in a place of its occurrence inside the global type.
\end{itemize}

@@

\paragraph{Examples of definition of global types}
Type generated from element named "GlobalElement", containing elements "Text", which is string, and element "AnotherGlobalElement", which is another global element.

@@

\paragraph{RootElementExporter}
\code{RootElementExporter} retrieves the root element from the result of preprocessing and creates its definition. This definition is also appended to the \code{indentator} (supplied to constructor) at a level of indentation, the \code{indentator} is set. Way of handling the root element's subelements depends on if these are global elements or not.\\

@@

Elements which are not global and are not of a built-in type are defined recursively, at the place of their occurrence. For example, root element "Root" contains element "A", which contains two other string elements.

@@

Export of attributes is done alongside with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.

@@

\begin{itemize}
	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of the resulting XSD. See \ref{section:problems}.
	\item Minimal number of occurrences of element to define its type globally. (Applies only if generation of global elements is active.)
	\item Number of spaces in output per one level of indentation.
	\item Global type name prefix. It is a string which will be inserted before a name of a type, which is derived from element's name. Can be also an empty string. (Applies only if generation of global elements is active.)
	\item Global type name suffix. It is a string which will be appended after a name of a type, which is derived from element's name. Can be also an empty string. (Applies only if generation of global elements is active.)
\end{itemize}

@@

This module is just basic implementation, thus many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.

@@

This problem appears if input grammar contains an alternation of two or more concatenations, which share the same prefix (begin with at least one same element). An example of offending XSD follows.

\begin{verbatim}
...
  <xs:choice>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="B" type="TB"/>
      <xs:element name="C" type="TC"/>
    </xs:sequence>
    <xs:sequence>
      <xs:element name="A" type="TA"/>
      <xs:element name="B" type="TB"/>
      <xs:element name="D" type="TD"/>
      <xs:element name="E" type="TE"/>
    </xs:sequence>
  </choice>
...
\end{verbatim}

@@

To make the result valid, manual modification is needed. The same prefix elements have to be removed from both sequences and inserted into a new \code{<xs:sequence>} element before the \code{<xs:choice>} element.\\

Adjusted XSD from the example can look like the following.
\begin{verbatim}
...
  <xs:sequence>
    <xs:element name="A" type="TA"/>
    <xs:element name="B" type="TB"/>
  <xs:sequence>
  <xs:choice>
    <xs:element name="C" type="TC"/>
    <xs:sequence>
      <xs:element name="D" type="TD"/>
      <xs:element name="E" type="TE"/>
    </xs:sequence>
  </choice>
...
\end{verbatim}

@@

Flow of data in this module is the following.
\begin{enumerate}
	\item \code{Preprocessor} topologically sorts elements (rules) it got on input.
	\item For each element, its occurrence count is computed.
	\item Unused elements (occurrence count equals 0) are removed.
	\item For each element, it is determined if it type will be defined as a global type or not.
	\item For each global element, its type is exported to a XSD representation by \code{GlobalElementsExporter}.
	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{RootElementExporter}.
	\item String representation of the schema is returned along with the information that file extension should be "xsd".
\end{enumerate}
